manuscript

### 2021-03-23 reports/manuscript.Rmd
Udpates results knitrs

We have observed that patients suffering from schizophrenia had significantly worse baseline RBANS performance than patietns with affective disorders (see table XXX). This was one of the reasons we opted for the linear mixed modelling in our subsequent analyses as they should be able to control for the effect of diagnosis.

@@

lmer_time_session_trial <- lmer(results_time ~ session + n_items + (1|participant), data = df_results_remediation)
lmer_trajectory_session_trial <- lmer(results_trajectory ~ session + n_items + (1|participant), data = df_results_remediation)
lmer_correct_session_6_9 <- lmer(correct_ratio ~ session + n_items + (1|participant), data = filter(df_results_remediation, n_items >= 6))

lmer_ratio_session <- df_results_remediation %>%
  lmer(correct_ratio ~ session + n_items + (1|rbansid), data = .)
summary(lmer_ratio_session)
```

@@

```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(fac_subtest, score, fill = trenink)) +
    geom_boxplot() +
    facet_wrap(~session,  strip.position = "bottom") +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
      "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
      "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme_apa() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
    labs(x = "", y = "RBANS IQ score",
         title = "IQ scores of RBANS cognitive domains",
         fill = "")

@@

@@

```{r session model explanation,eval = FALSE}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname) %>%
    lmer(z_score ~ factor(session) + (1|rbansid), data = .)
  return(out)
}

@@

```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(subtest, score, fill = trenink)) +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
    "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
    "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    geom_boxplot() + 
    theme_apa() +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
      labs(x = "", y = "RBANS IQ score",
         title = "IQ scores of RBANS cognitive domains after each intervention",
         fill = "")

@@

```{r}
temp_mixed_func <- function(varname){
  out <- df_temp %>%
    filter(subtest == varname, session > 1) %>%
    lmer(score ~ total_score + learning_baseline + factor(session) + (1|rbansid), data = .)
  return(out)
}

>>>>

We have observed that patients suffering from schizophrenia had significantly worse baseline RBANS performance than patients with MDD (see table XXX). This was one of the reasons we opted for the linear mixed modeling in our subsequent analyses as they should be able to control for the effect of diagnosis.

@@

lmer_time_session_trial <- lmer(results_time ~ session + n_items + (1|participant), 
                                data = df_results_remediation)
lmer_trajectory_session_trial <- lmer(results_trajectory ~ session + n_items + (1|participant), 
                                      data = df_results_remediation)
lmer_correct_session_6_9 <- lmer(correct_ratio ~ session + n_items + (1|participant), 
                                 data = filter(df_results_remediation, n_items >= 6))
```

@@

```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(fac_subtest, score, fill = trenink)) +
    geom_boxplot() +
    facet_wrap(~session,  strip.position = "bottom") +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
      "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
      "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme_apa() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
    labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains",
         fill = "")

@@

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 2)})
knitr::kable(as.data.frame(res),
             caption = "Trenink as a predictor of RBANS performance. 
             Coefficients are for VR in contrast to TP (negative = VR does worse)")
```

@@

```{r session model explanation, eval = FALSE}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname) %>%
    lmer(z_score ~ factor(session) + (1|rbansid), data = .)
  return(out)
}

@@

## Supplementary

```{r diagnosis model, eval = TRUE}
temp_mixed_func <- function(varname){
  out <- df_rbans_long %>%
    filter(subtest == varname, session > 1) %>%
    lmer(z_score ~ trenink + diagnosis + factor(session) + (1|rbansid), data = .)
  return(out)
}

dims <- c("dim_celek", "dim_kratkodobapamet", "dim_dlouhodobapamet",
          "dim_pozornost", "dim_rec", "dim_vizuoprostor")
res <- sapply(dims, function(x){report_mixed_effect(temp_mixed_func(x), 3)})
knitr::kable(as.data.frame(res),
             caption = "Diagnosis as a predictor of RBANS performance. 
             Coefficients are for schizophrenia in contrast to MDD (negative = schizophrenia does worse)")
```
```{r}
df_rbans_long %>%
  filter(grepl("iq_", subtest)) %>%
  mutate(session = paste0("session ", session - 1),
         fac_subtest = factor(subtest, levels = c("iq_celek", "iq_kratkodobapamet",
                                                  "iq_dlouhodobapamet", "iq_rec",
                                                  "iq_pozornost", "iq_vizuoprostorschopnosti"))) %>%
  ggplot(aes(subtest, score, fill = trenink)) +
    scale_x_discrete(labels = c("iq_celek" = "Total", "iq_kratkodobapamet" = "Immediate memory", 
    "iq_vizuoprostorschopnosti" = "Visuospatial/constructional", "iq_rec" = "Language", 
    "iq_pozornost" = "Attention", "iq_dlouhodobapamet" = "Delayed memory")) +
    geom_boxplot() + 
    theme_apa() +
    scale_fill_manual(labels = c("baseline", "TP" = "Standard treatment",
                                   "VR" = "VEs program"),
                        values = CDZ_COLORS) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1),
          legend.position = c(0.5,1), legend.direction = "horizontal") +
      labs(x = "", y = "RBANS index score",
         title = "Index scores of RBANS cognitive domains after each intervention",
         fill = "")

@@

```{r effect of testuceni, eval=FALSE}
temp_mixed_func <- function(varname){
  out <- df_temp %>%
    filter(subtest == varname, session > 1) %>%
    lmer(score ~ total_score + learning_baseline + factor(session) + (1|rbansid), data = .)
  return(out)
}
